---
title: "exp_spatially_dependent_item_gen"
author: "Gabriel Strain"
date: '2022-05-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in packages

```{r}
library(MASS)
library(tidyverse)
library(Hmisc)
library(modelr)
library(scales)
library(gtools)
```

# Generating correlations

```{r}
# We only want one unique set of plots

number_of_participants = 1

# With 4 levels of IV1 we 180/4 = 45 trials

number_of_trials = 45

set.seed(1234)

my_ss <- rep(1:number_of_participants, each = number_of_trials)

my_rs <- seq(0.2, 0.99, length.out = 45)

correlations <- tibble(my_ss = rep(my_ss, each = 4),  my_rs = rep(my_rs, each = 4))
```

# Creating the residuals function

## This function will add the sum of residuals for each plot

```{r}
add_residual_function <- function(my_desired_r) {
  
  set.seed(1234)

  my_sample_size = 128
  
  mean_variable_1 = 0
  sd_variable_1 = 1
  
  mean_variable_2 = 0
  sd_variable_2 = 1
  
  mu <- c(mean_variable_1, mean_variable_2) 
  
  myr <- my_desired_r * sqrt(sd_variable_1) * sqrt(sd_variable_2)
  
  mysigma <- matrix(c(sd_variable_1, myr, myr, sd_variable_2), 2, 2) 
  
  corr_data = as_tibble(mvrnorm(my_sample_size, mu, mysigma, empirical = TRUE))
  
  corr_model <- lm(V2 ~ V1, data = corr_data)
  
  my_residuals <- abs(residuals(corr_model))
  
  data_with_resid <- round(cbind(corr_data, my_residuals), 2)
  
  return(sum(data_with_resid$my_residuals))
  
}
```

# Creating necessary components for loop

## Here we create somewhere to store the sum residuals (r_vec)

```{r}
r_vec <- correlations$my_rs
total_residuals <- c()
```

# Writing loop to add total residuals

```{r}
set.seed(1234)

for(value in r_vec) {
  new_value <- add_residual_function(value)
  total_residuals <- c(total_residuals, new_value) 
}
```

# Concatenating correlation data and residuals 

```{r}
item_no <- rep(1:45, each = 4)
full_data <- cbind(item_no, correlations, total_residuals)
```

# Checking the max residual for a particular seed - we need this for linear scaling

```{r}
set.seed(1234)

corr_values <- seq(0.2, 0.99, length.out = 45)

max_residual <- function(my_desired_r){
  
my_sample_size = 128

mean_variable_1 = 0
sd_variable_1 = 1

mean_variable_2 = 0
sd_variable_2 = 1

mu <- c(mean_variable_1, mean_variable_2) 

myr <- my_desired_r * sqrt(sd_variable_1) * sqrt(sd_variable_2)

mysigma <- matrix(c(sd_variable_1, myr, myr, sd_variable_2), 2, 2) 

corr_data = as_tibble(mvrnorm(my_sample_size, mu, mysigma, empirical = TRUE))

corr_model <- lm(V2 ~ V1, data = corr_data)

my_residuals <- abs(residuals(corr_model))

data_with_resid <- round(cbind(corr_data, my_residuals), 2)

return(max(abs(data_with_resid$my_residuals)))

}

c <- as_tibble(sapply(corr_values, max_residual))

max(c$value)
```

# Slope generation function

```{r}
# A: Non-Linear Decay, a = 0.25
# B: Linear Decay
# C: Non-Linear Decay, Inverted
# D: Full Contrast
# This function generates the slope values needed to create plots with the various decay parameters.
# Maximum residual with this seed is 3.2, so for linear slope, we divide all residuals by 3.2

slope_function <- function(my_desired_r) {
  set.seed(1234)
  
  my_sample_size = 128
  
  mean_variable_1 = 0
  sd_variable_1 = 1
  
  mean_variable_2 = 0
  sd_variable_2 = 1
  
  mu <- c(mean_variable_1, mean_variable_2) 
  
  myr <- my_desired_r * sqrt(sd_variable_1) * sqrt(sd_variable_2)
  
  mysigma <- matrix(c(sd_variable_1, myr, myr, sd_variable_2), 2, 2) 
  
  corr_data = as_tibble(mvrnorm(my_sample_size, mu, mysigma, empirical = TRUE))
  
  corr_model <- lm(V2 ~ V1, data = corr_data)
  
  my_residuals <- abs(residuals(corr_model))
  
  data_with_resid <- round(cbind(corr_data, my_residuals), 2)
  
slopes <- data_with_resid %>%
  mutate(slope_linear = my_residuals/3.2) %>%
  mutate(slope_0.25 = 1-(0.25)^my_residuals) %>%
  mutate(slope_inverted = (1 + (0.25)^ my_residuals)-1)

return(slopes)
}
```
 
# Plot generation function

```{r}
plot_function <- function(slopes, my_desired_r, letter, contrast) {
  
  p <- ggplot(slopes, aes(x = V1, y = V2)) +
  scale_alpha_identity() +
  geom_point(aes(alpha = contrast), shape = 16) +
  labs(x = "", y = "") +
  theme_classic() +
  theme(axis.text = element_blank()) +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(legend.position = "None")

ggsave(p,filename=paste0(counter, letter,".png"),
       device = "png",
       bg = "white",
       path = "all_plots",
       units = "px",
       width = 1200,
       height = 1200,
)
}
```

# Creating the plots

```{r}
# This loop iterates the function above over the 45 r values in my_rs

# Running this chunk will use the function we defined above on each r value, 
# then saves the plot with the proper labeling

contrastlevels <- list()


abcd <- c('A','B','C','D')

for (i in abcd) {
  counter = 1
  for (value in my_rs) {
    slopes <- slope_function(value)
    
contrastlevels[['A']] <- (1-(slopes$slope_0.25))
contrastlevels[['B']] <- (1-(slopes$slope_linear))
contrastlevels[['C']] <- (1-(slopes$slope_inverted))
contrastlevels[['D']] <- 1

    plot_function(slopes, value,i,contrastlevels[[i]])
    if (counter > 0) {
      counter = counter + 1
    }
  }
}

```

# Assembling .csv with plot file names, unique item numbers, r values, residuals

```{r}
images <- mixedsort(list.files(path = "all_plots"))

plot_labels <- rep(c("all_plots/"), each = 180)

plots_with_labels <- paste(plot_labels, images, sep = "")

final_data <- full_data %>%
  select(-c(item_no, my_ss))

unique_item_no <- c(1:180)

instructions <- rep(c("Please look at the following plot and use the slider to estimate the correlation"), each = 180)

data_with_plots <- cbind(unique_item_no, final_data, plots_with_labels, images, instructions)
```

# Creating masks

```{r}
mask_gen_function <- function() {

x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- round(runif(400, 0, 1))

mask <- ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_gradient(low = "white", high = "black") +
  theme_void() +
  theme(legend.position = "None")

ggsave(mask,filename=paste("mask", counter,".png",sep=""),
       device = "png",
       bg = "white",
       path = "masks",
       units = "px",
       width = 1200,
       height = 1200,)
}

counter = 1
for (i in 1:186) {
  mask_gen_function()
  if (counter > 0) {
    counter = counter + 1
  }
}

masks <- mixedsort(list.files(path = "masks"))

mask_labels <- rep(c("masks/"), each = 186)

masks_with_labels <- paste(mask_labels, masks, sep = "")
```

# Attaching height and width values to each plot, instructions column, and masks vector

```{r}
size_w <- rep(c(0.86), times = 180)

size_h <- rep(c(0.86), times = 180)

colour <- rep(c("black"), times = 180)

data_with_sizes <- cbind(data_with_plots, size_h, size_w, colour)
```

# Creating attention_check DF.

```{r}
my_rs <- rep(c(0, 1), times = c(3, 3))

total_residuals <- rep(c(0), times = 6)

plots_with_labels <- head(plots_with_labels)

images <- head(images)

unique_item_no <- c(181:186)

size_h <- rep(c(0.86), times = 6)

size_w <- rep(c(0.86), times = 6)

colour <- rep(c("red"), times = 6)

instructions <- rep(c("Please IGNORE the correlation displayed and set the slider to 0.",
                      "Please IGNORE the correlation displayed and set the slider to 1."),
                    times = c(3, 3))

attention_checks <- cbind(unique_item_no, my_rs, total_residuals, plots_with_labels, images,
                          size_h, size_w, instructions, colour)
```

# Creating the final DF by rbinding attention_checks with data_with_sizes

```{r}
exp_spatially_dependent_final <- rbind(data_with_sizes, attention_checks) %>%
  cbind(masks_with_labels)

write.csv(exp_spatially_dependent_final, "exp_spatially_dependent_full_data.csv", row.names = FALSE)
```

# Practice trials

## Prepare practice trial data

```{r}
number_of_participants = 1

number_of_trials = 2

set.seed(12345)

my_ss <- rep(1:number_of_participants, each = number_of_trials)

my_rs <- rep(runif(number_of_trials, min = 0.2, max = 0.9), number_of_participants)

correlations <- tibble(my_ss, my_rs)
```

# Practice trial generation function

```{r}
practice_plot_generation_function <- function(my_desired_r) {
  set.seed(12345)
  
  my_sample_size = 128
  
  mean_variable_1 = 0
  sd_variable_1 = 1
  
  mean_variable_2 = 0
  sd_variable_2 = 1
  
  mu <- c(mean_variable_1, mean_variable_2) 
  
  myr <- my_desired_r * sqrt(sd_variable_1) * sqrt(sd_variable_2)
  
  mysigma <- matrix(c(sd_variable_1, myr, myr, sd_variable_2), 2, 2) 
  
  corr_data = as_tibble(mvrnorm(my_sample_size, mu, mysigma, empirical = TRUE))
  
  corr_model <- lm(V2 ~ V1, data = corr_data)
  
  my_residuals <- abs(residuals(corr_model))
  
  data_with_resid <- round(cbind(corr_data, my_residuals), 2)

test_plot <- ggplot(data_with_resid, aes(x = V1, y = V2)) +
  geom_point()  +
  labs(x = "", y = "") +
  theme_classic() +
  theme(axis.text = element_blank()) +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(legend.position = "None")

ggsave(test_plot,filename=paste0(counter,"P",".png"), 
       device = "png",
       bg = "white",
       path = "practice_plots",
       units = "px",
       width = 1200,
       height = 1200,
)

return()
}
```

# Creating practice plots

```{r}
counter = 1
for (value in my_rs) {
  practice_plot_generation_function(value)
  if (counter > 0) {
    counter = counter + 1
  }
  }
```

# Arranging practice plots into new .csv

```{r}
practice_images <- mixedsort(list.files(path = "practice_plots"))

size_h_p <- rep(c(0.86), times = 2)

size_w_p <- rep(c(0.86), times = 2)

practice_data <- cbind(practice_images, size_h_p, size_w_p)

write.csv(practice_data, "practice_plots.csv", row.names = FALSE)
```
